---
title: "Data Import"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data.

## Import the FAS Litters CSV
clean_names() replaces spaces with underscores, removes special characters, etc.
Also, the janitor::clean_names() function pulls this function without loading the entire janitor package
```{r}
litters_df = read_csv("data_import_examples/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df) 
```

## Look at the dataset
```{r}
litters_df
head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
view(litters_df)
```

## Learning check: import FAS pups csv using relative path
```{r}
pups_df = read_csv("data_import_examples/FAS_pups.csv")
```

## import FAS pups csv using absolute path
```{r eval = FALSE}
pups_df = read_csv("/Users/dylanmorgan/Desktop/Data_Sci_I/data_wrangling_i/data_import_examples/FAS_pups.csv")
```

## Look at read_csv options
```{r}
litters_df = 
  read_csv(
    file = "./data_import_examples/FAS_litters.csv", 
    col_names = FALSE, 
    skip = 1
  )
```

What about missing data?
```{r}
litters_df = 
  read_csv(
    file = "./data_import_examples/FAS_litters.csv", 
    na = c(".", "NA", "")
  )

litters_df = janitor::clean_names(litters_df)
pull(litters_df, gd0_weight)
```

What if we code `group` as a factor variable?
```{r}
litters_df = 
  read_csv(
    file = "data_import_examples/FAS_litters.csv", 
    na = c(".", "NA", ""), 
    col_types = cols(
      Group = col_factor()
    )
  )
```

## Import an excel file
Import MLB 2011 summary data.
```{r}
mlb_df = read_excel("data_import_examples/mlb11.xlsx", sheet = "mlb11")
```

## Import SAS data
```{r}
pulse_df = read_sas("data_import_examples/public_pulse_data.sas7bdat")
```

## Never use read.csv()
read.csv prints entire dataset, whereas read_csv will only print first 10 lines (better for memory)
```{r}
litters_df = read.csv("data_import_examples/FAS_litters.csv")
```

Never do this either because this involves taking data out of the dataset
```{r}
litters_df$L
```





